---
title: "DATA 605 Assignment 9"
author: "Warner Alexis"
date: "2024-03-25"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Excercise 11 page 363
A tourist in Las Vegas was attracted by a certain gambling game in which
the customer stakes 1 dollar on each play; a win then pays the customer
2 dollars plus the return of her stake, although a loss costs her only her stake.
Las Vegas insiders, and alert students of probability theory, know that the
probability of winning at this game is 1/4. When driven from the tables by
hunger, the tourist had played this game 240 times. Assuming that no near
miracles happened, about how much poorer was the tourist upon leaving the
casino? What is the probability that she lost no money?


## Solution 


Lets $X_n = Y_n + 1$ $Y_n$ is an independent  random variables having a mean $\mu = 0$
```{r}
sd <- sqrt(365*.25) #will use this for sd for a, b, & c
q <- 100-100
pr  <- pnorm(q, 0, sd, lower.tail = FALSE)
cat("the probability that Y365 is ≥ 100: ", pr, "\n")


# calculating x as : (value - mean)/sqrt(n)
mean <- 0
variance <- 1/4
sd <- sqrt(variance)
n <- 364

q2 <- (110 - 100)/sqrt(n)
cat("the probability that Y365 is ≥ 110: ",pnorm(q2, mean = mean, sd = sd, lower.tail = FALSE),"\n")


## calculating x as : (value - mean)/sqrt(n)
q3 <- (120 - 100)/sqrt(n)
cat("the probability that Y365 is ≥ 120: ",pnorm(q3, mean = mean, sd = sd, lower.tail = FALSE),"\n")
```
## Question 2 
Moment generating function $Mz(t) = Ee^{tx}f(x)$ 
Binomial distribution function: 
$P(x) = \sum_{j=0} \dbinom{n}{x}p^xq^{(n-x)}$
$P(x) = \sum_{j=0} \dbinom{n}{x}(pe^t)q^{(n-x)}$
$P(x) = (pe^t)q)^n$
$g'(t) = n(pe^t + q)^{n-1}pe^t$
$g'(0) = n(p + q)^{n-1}p$
$g'(0) = np$
                                
   then, 
                                $g′′(0)=n(n−1)p^2+np$
                                $Var=n(n−1)p^2+np−(np)^2$
                                $=(n^2−n)(p^2)+np−n^2p^2$
                                $=np(1−p)$
                                
## Question 3
Calculate the expected value and variance of the exponential distribution using the moment generating function.
                                $G(t) = \int_{0}^\infty e^{tx} \lambda e^{- \lambda x}dx$
                                $\frac {\lambda(e^{(t- \lambda)} }{t- \lambda} \vert_0 ^\infty$
                                $\frac {\lambda}{\lambda - t}$
                                $g′(t) = \frac {\lambda}{(\lambda - t)^2}$
                                $g′(0) = \frac {\lambda}{(\lambda)^2}$
                                $g′(0) = \frac {1}{\lambda}$
                                
The second derivative evaluated at 0, and the square of the first derivative evaluated at 0. 
                                $g"(t) = \frac {2\lambda}{(\lambda - t)^3}$
                                $g′(0) = \frac {2\lambda}{(\lambda)^3}$
                                $g′(0) = \frac {2}{(\lambda)^2}$
                                $Var = \frac {2}{(\lambda)^2} - ( \frac {1}{\lambda} )^2$
                                $Var = \frac {2}{\lambda^2} -  \frac {1}{\lambda}$
                                $Var = \frac {1}{\lambda^2}$
                                
he variance of an exponential distribution is $\frac {1}{\lambda^2}$
                                
