---
title: "Ass 3"
author: "Warner Alexis"
date: "2024-02-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Assignment 3 
we have a matric A 
$A = \begin{bmatrix}1 & 2 & 3 & 4\\ -1 &0 & 1 & 3 \\ 0 & 1 & -2 & 1 \\ 5 & 4 & -2 & -3\end{bmatrix}$.
We are goijng to reduce the matrix to echelon form 
R2 <- R2 + R1
R4 <- R4 - 5R1

that will give us this matrix 
$A = \begin{bmatrix}1 & 2 & 3 & 4\\ 0 &2 & 4 & 7 \\ 0 & 1 & -2 & 1 \\ 0 & -6 & -17 & -23\end{bmatrix}$

we continue to break down the matrix so we can get the non zero rows 


R2 <- 1/2R2


R4 <- R4 - 6R3

That will give us 
$A = \begin{bmatrix}1 & 2 & 3 & 4\\ 0 &1 & 2 & 7/2 \\ 0 & 0 & -4 & 5/2 \\ 0 & 0 & -5 & -17\end{bmatrix}$

R3 <- R3 + R4


R4 <- R4 + R4

Final matrix is 
$A = \begin{bmatrix}1 & 2 & 3 & 4\\ 0 &1 & 2 & 7/2 \\ 0 & 0 & 1 & 24/2 \\ 0 & 0 & 1 & -39/2\end{bmatrix}$

Therefore, n,r(A) = n . we imply that rank is 4

```{r}
# initialize matrix
A <- matrix(c(1,2,3,4,-1,0,1,3,0,1,-2,1,5,4,-2,-3),nrow=4,byrow= TRUE)
A

# use matrix library 
library(Matrix)
rankMatrix(A)[1][1]



```



##2  Given an m x n matrix where m > n, what can be the maximum rank? The minimum rank, assuming that the matrix is non-zero?

Thee maximum rank a matrix m x n can have the maximum rank of n because it is not possible to have more than n linearly independent columns. The minimum rank assuming the matrix is non-zero would be 1 

#3 the rank for matrix 

$B = \begin{bmatrix}1 & 2 & 1\\ 3 &6 & 3  \\ 2 & 4 & 2 \end{bmatrix}$

We are going to reduce the matrix to echelon form 


R2 <- R2 -3R1


R3 <- R3 - 2R1

We will have this matrix: 

$B = \begin{bmatrix}1 & 2 & 1\\ 0 &0 & 0  \\ 0 & 0 & 0 \end{bmatrix}$

Therefore, n,r(A) = n . we imply that rank is 1


```{r}
B <- matrix(c(1,2,1,3,6,3,2,4,2), nrow = 3, ncol = 3, byrow = TRUE)
B
rankMatrix(B)[1][1]

```



## Q3

Lets a matrix 
$B = \begin{bmatrix}1 & 2 & 3\\ 0 &4 & 5  \\ 0 & 0 & 6 \end{bmatrix}$

find the eigenvalues and eigenvectors: 

## Step 1  

$|A - \lambda I | = 0$

so we have :
 $B = \begin{bmatrix}1 & 2 & 3\\ 0 &4 & 5  \\ 0 & 0 & 6 \end{bmatrix} * \lambda \begin{bmatrix}1 & 0 & 0\\ 0 &1 & 0  \\ 0 & 0 & 1 \end{bmatrix}  $
 
$B = \begin{bmatrix}1-\lambda & 2 & 3\\ 0 &4-\lambda & 5  \\ 0 & 0 & 6-\lambda \end{bmatrix}$ 

## Step 2 

$1 - \lambda \begin{bmatrix}1-\lambda & 5 \\ 0 &6-\lambda \end{bmatrix} - 2\begin{bmatrix}0 & 5 \\ 0 &6-\lambda \end{bmatrix} +3\begin{bmatrix}0 & 4-\lambda \\ 0 &0 \end{bmatrix}$

$(1 - \lambda) (4-\lambda) (6-\lambda)$

Eigenvalue will be 6 4 1 

$(1 - \lambda) (4-\lambda) (6-\lambda)$
$(\lambda^3 - 5\lambda^2 +4\lambda)-(6\lambda^2 -30\lambda + 24)$
$\lambda^3 - 11\lambda^2 - 34\lambda - 24$

```{r}
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow = 3, ncol = 3, byrow = TRUE)
eigen(A)

```




Calculate the eigenvector.
we find the eigenvectors corresponding to each eigenvalue by solving the equation $(A-\lambda)v =0 $

for $(\lambda - 6)$ :


                   
$B = \begin{bmatrix}1-6 & 2 & 3\\ 0 &4-6 & 5  \\ 0 & 0 & 6-6 \end{bmatrix}$

$R1 <- -1/5R1$
$R2 <- -1/5R2$

$B = \begin{bmatrix}1 & -2/5 & -3/5\\ 0 &1 & 5/2  \\ 0 & 0 & 0 \end{bmatrix}$

$v2 - 5/2v3 = 0$
$v1 - 2/5v2-3/5v3 = 0$
$v1 = t$

for $(\lambda - 6) =  \begin{bmatrix}t & 5/2 & t \end{bmatrix}$


for $(\lambda - 4)$ :


                   
$A = \begin{bmatrix}1-4 & 2 & 3\\ 0 &4-4 & 5  \\ 0 & 0 & 6-4 \end{bmatrix}$

$R1 <- -1/3R1$


$B = \begin{bmatrix}1 & -2/53 & -1\\ 0 &0 & 5  \\ 0 & 0 & 2 \end{bmatrix}$

$5v3 = 0$
$v1 - 2/3v2-v3 = 0$
$v1 = -2/3t$


for $(\lambda - 4) =  \begin{bmatrix}-2/3 & 1 & 0 \end{bmatrix}$

```{r}
# calculate the eigenvector
A <- matrix(c(1,2,3,0,4,5,0,0,6), nrow = 3, ncol = 3, byrow = TRUE)

eigen(A)[2]
```



