---
title: "Data 605 Assignment 10"
author: "Warner Alexis"
date: "2024-04-01"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Excercise

Smith is in jail and has 1 dollar; he can get out on bail if he has 8 dollars.
A guard agrees to make a series of bets with him. If Smith bets A dollars,
he wins A dollars with probability .4 and loses A dollars with probability .6.
Find the probability that he wins 8 dollars before losing all of his money if
(a) he bets 1 dollar each time (timid strategy).
(b) he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).
(c) Which strategy gives Smith the better chance of getting out of jail?

## solution 

Lets represent Smith fortunes: 
State 0: Smith has lost all his money.
State 1: Smith has 1 dollar.
State 2: Smith has 2 dollars.
State 8: Smith has 8 dollars and can get out on bail.

The transition probabilities between states depend on whether Smith wins or loses the bet. Let 
p be the probability of winning a bet (0.4), and
q be the probability of losing a bet (0.6).

```{r}

library(markovchain)
outcomes <- c('0','1','2','3','4','5','6','7','8') 


# Define transition matrix for the timid strategy
trans_matrix_1 = matrix(c(1,0,0,0,0,0,0,0,0,
                          0.6,0,0.4,0,0,0,0,0,0,
                          0,0.6,0,0.4,0,0,0,0,0,
                          0,0,0.6,0,0.4,0,0,0,0,
                          0,0,0,0.6,0,0.4,0,0,0,
                          0,0,0,0,0.6,0,0.4,0,0,
                          0,0,0,0,0,0.6,0,0.4,0,
                          0,0,0,0,0,0,0.6,0,0.4,
                          0,0,0,0,0,0,0,0,1), 
                        byrow = T, nrow = 9, 
                        dimnames = list(outcomes,outcomes))
trans_matrix_1

# Create markovchain objects
mc_timid <- new("markovchain", transitionMatrix = trans_matrix_1)
absorptionProbabilities(mc_timid)
```


The proability when he bet each is time is 0.02030135

b)  he bets, each time, as much as possible but not more than necessary to
bring his fortune up to 8 dollars (bold strategy).

```{r}
trans_matrix_2 = matrix(c(1,0,0,0,0,0,0,0,0,
                          0.6,0,0.4,0,0,0,0,0,0,
                          0.6,0,0,0,0.4,0,0,0,0,
                          0.6,0,0,0,0,0,0.4,0,0,
                          0.6,0,0,0,0,0,0,0,0.4,
                          0,0,0.6,0,0,0,0,0,0.4,
                          0,0,0,0,0.6,0,0,0,0.4,
                          0,0,0,0,0,0,0.6,0,0.4,
                          0,0,0,0,0,0,0,0,1), 
                        byrow = T, nrow = 9, 
                        dimnames = list(outcomes,outcomes))
mc_bold <- new("markovchain", transitionMatrix = trans_matrix_2)
absorptionProbabilities(mc_bold)
```

The probability that he brings his fortune up to 8 dollars is 0.064 on his first try. 


(c) Which strategy gives Smith the better chance of getting out of jail?

The bold tratejy give Smith a better chance of getting out of jail. bold is 6.04%  compare to 2.03%.
