---
title: "Data 605 Assignment Week 8"
author: "Warner Alexis"
date: "2024-03-18"
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## page 303  number 11

A company buys 100 lightbulbs, each of which has an exponential lifetime of
1000 hours. What is the expected time for the rst of these bulbs to burn
out? (See Exercise 10.)

**Solution**

Lets $(X_i - X_n)$ be independent variables with parameters $\lambda_i ...\lambda_n $
$Pr(k|X_k = minX_i...X_n) = \frac {\lambda_k}{\lambda_i ...\lambda_n}$
$\lambda_i = \frac {1}{1000}, \sum_{\lambda_i} =\frac{1}{10} $
$\mu = 1000, n = 100$

```{r}
mu = 1000
bulb_n = 100

cat("The Expected time for the first of these bulbs to burn out is ", mu/bulb_n ,"hours ")
```
## page 303 # 14

Particles are subject to collisions that cause them to split into two parts with
each part a fraction of the parent. Suppose that this fraction is uniformly
distributed between 0 and 1. Following a single particle through several splittings
we obtain a fraction of the original particle $Zn = X_1*X_2...X_n$ where
each Xj is uniformly distributed between 0 and 1. Show that the density for
the random variable $Z_n$ is

$fn^(z^) = \frac {1}{(n-1)!}(-log z)^n-1$

**Solution**

Lets Z = X1 - X2 when z > 0 
                            $z = x_1 - x_2$
                            $x_2 = x_1 - z$
                            $fZ(x) = \int_{\infty}^\infty f_x1(x_1) f_x2(x1 - z)dx_1$
                            Solve it by exponential distribution
                            $fX(x) = \lambda e ^{-\lambda x}$
                            $\int_{0}^\infty \lambda ^{2\lambda x_1} \lambda e ^{-\lambda (x_1 - z)} dx_1$
                            $\int_{0}^\infty \lambda ^2e^{\lambda( -2x_1 + \lambda z)}  dx_1$
                            $\int_{0}^\infty \lambda ^2e^{\lambda (z-2x_1)}dx_1$
                            $(1/2)\lambda 2^{-\lambda(z)}$
                            
##   page 320-321   

A fair coin is tossed 100 times. The expected number of heads is 50, and the
standard deviation for the number of heads is (100  1=2  1=2)1=2 = 5. What
does Chebyshev's Inequality tell you about the probability that the number
of heads that turn up deviates from the expected number 50 by three or more
standard deviations (i.e., by at least 15)?

**Solution**

Suppose that we have the following input:
                          $\mu = 10 , variance = 100/3$
                          $P(|X-\mu|)  >= k\sigma \le \frac {\sigma^2}{k^2\sigma^2} = \frac {1}{k^2}$
                          $P(|X-10|)  >= 2 \le \frac {1}{2^2} $ 
                          
```{r}
#MEan 
mu = 10 
# variance
var = 100/3
# Standard deviation 
sd = sqrt(var)
ksd = 2 
k <- ksd / sd

upper_bnd = 1/(k^2)


cat("The the upper bound is ", upper_bnd,"\n")
cat("The propability of the upper bound is ", pmin(upper_bnd))



```
  
When $P(|X-10|)  >+ 5$ then we will have : 
                          $P(|X-10|)  >= 5 \le \frac {1}{5^2} $
                          
```{r}
sd_k  = 5 

k <- 5 / sd
upper_bnd = 1/(k^2)
cat("The highest propability is  ", upper_bnd,"\n")
cat("The propability of the upper bound is ", pmin(upper_bnd))

```
                        
When $P(|X-10|)  >= 9$ then we will have : 
                          $P(|X-10|)  >= 9 \le \frac {1}{9^2} $
                          
```{r}
sd_k  = 9

k <- sd_k / sd
upper_bnd = 1/(k^2)
cat("The highest propability is  ", upper_bnd,"\n")
cat("The propability of the upper bound is ", pmin(upper_bnd))
                          
```
                          
 When $P(|X-10|)  >= 20$ then we will have : 
                          $P(|X-10|)  >= 20 \le \frac {1}{20^2} $     
                          
                          
```{r}
sd_k  = 20 

k <- sd_k / sd
upper_bnd = 1/(k^2)
cat("The highest propability is  ", upper_bnd,"\n")
cat("The propability of the upper bound is ", pmin(upper_bnd))
```
                          
                            




