---
title: "data"
author: "Warner Alexis"
date: "2024-04-08"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## dna analysis


We are importing two data sets from Kaggle using DNA data set test and training sets. x and y variables seems to be well correlated so we expect the $R2$ to be close to 1.





```{r}
library(ggplot2) 
library(dplyr)

test <- read.csv('https://raw.githubusercontent.com/joewarner89/DATA-605-Computational-Mathematics/main/Assignment/week%2012/test.csv')

train <- read.csv('https://raw.githubusercontent.com/joewarner89/DATA-605-Computational-Mathematics/main/Assignment/week%2012/train.csv')
# Check for NA and missing values
# is.na return a vector with value TT for missing values.
numberOfNA = length(which(is.na(train)==T))
if(numberOfNA > 0) {
  cat('Number of missing values found: ', numberOfNA)
  cat('\nRemoving missing values...')
  train = train[complete.cases(train), ]
}

cor(train$x,train$y)
```





## Simple Linear Regression

Both dna strands x and y are well correlated. x and y are well distributed.


```{r}


require(ResourceSelection)
#summarize the data set
kdepairs(train) 
#creating the model 
dna_lm <- lm(y~., data = train) 
summary(dna_lm)
```







```{r}
par(ask=F)
par(mfrow=c(2,2))
plot(dna_lm)
hist(dna_lm$residuals)

```




The model acheives a $R^2$ = 0.9907 and low p-value that are significant for the model development. we can use this model to predict chromosome y from x from the DNA simple. All four plots in the summary statistics of the model suggest a good model.

